SeedLing: Building and using a seed corpus \\ for the Human Language Project

### 2014-04-28 docs/computEL-final/concl.tex
Update number of languages and families.

In this paper, we have described the creation of SeedLing, a foundation text for a universal corpus, following the guidelines of Abney and Bird (2010; 2011). To do this, we cleaned and standardised data from several multilingual data sources: ODIN, Omniglot, the UDHR, Wikipedia. The resulting corpus is more easily machine-readable than any of the underlying data sources, and has been stored according to the best practices suggested by \newcite{bird2003port}. At present, SeedLing has data from 19\% of the world's languages, covering 72\% of language families. We believe that a corpus with such diversity of languages, uniformity of format, cleanliness of data, and ease of access provides an excellent seed for a universal corpus. It is our hope that taking steps toward creating this resource will spur both further data contributions and interesting computational research with cross-linguistic or typological perspectives; we have here demonstrated SeedLing's utility for such research by using the data to perform language clustering, with promising results.

SeedLing (both the data and the API) is currently available via a GitHub repository.\footnote{\url{https://github.com/alvations/SeedLing}} We have yet to fully address questions of long-term access, and we welcome ideas or collaborations along these lines.

>>>>

In this paper, we have described the creation of SeedLing, a foundation text for a universal corpus, following the guidelines of Abney and Bird (2010; 2011). To do this, we cleaned and standardised data from several multilingual data sources: ODIN, Omniglot, the UDHR, Wikipedia. The resulting corpus is more easily machine-readable than any of the underlying data sources, and has been stored according to the best practices suggested by \newcite{bird2003port}. At present, SeedLing has data from 19\% of the world's living languages, covering 72\% of language families. We believe that a corpus with such diversity of languages, uniformity of format, cleanliness of data, and ease of access provides an excellent seed for a universal corpus. It is our hope that taking steps toward creating this resource will spur both further data contributions and interesting computational research with cross-linguistic or typological perspectives; we have here demonstrated SeedLing's utility for such research by using the data to perform language clustering, with promising results.

SeedLing (data, API and documentation) is currently available via a GitHub repository.\footnote{\url{https://github.com/alvations/SeedLing}} We have yet to fully address questions of long-term access, and we welcome ideas or collaborations along these lines.

### 2014-04-28 docs/computEL-final/intro.tex
Update number of languages and families.

We believe the resulting corpus, which at present covers 1347 languages from 106 language families, is the first of its kind: large enough and consistent enough to allow broadly multilingual language processing. To test this claim, we use SeedLing in a sample application (section \ref{sec:cluster}): the task of language clustering. With no additional pre-processing required, we extract surface-level features (frequencies of character n-grams and words) for estimating the similarity of two languages. Unlike most previous approaches to the task, we make no use of resources curated for aims of linguistic typology (e.g. values of typological features as in WALS \cite{wals}, Swadesh word lists). Despite our approach being highly dependent on orthography, our clustering performance matches the results obtained by \newcite{georgi2010wals} using typolological features, which demonstrates SeedLing's utility in cross-linguistic research.

>>>>

We believe the resulting corpus, which at present covers 1451 languages from 105 language families, is the first of its kind: large enough and consistent enough to allow broadly multilingual language processing. To test this claim, we use SeedLing in a sample application (section \ref{sec:cluster}): the task of language clustering. With no additional pre-processing required, we extract surface-level features (frequencies of character n-grams and words) for estimating the similarity of two languages. Unlike most previous approaches to the task, we make no use of resources curated for aims of linguistic typology (e.g. values of typological features as in WALS \cite{wals}, Swadesh word lists). Despite our approach being highly dependent on orthography, our clustering performance matches the results obtained by \newcite{georgi2010wals} using typolological features, which demonstrates SeedLing's utility in cross-linguistic research.

### 2014-04-28 docs/computEL-final/paper.tex
Update number of languages and families.

\begin{document}
\maketitle
\begin{abstract}
  A broad-coverage corpus such as the Human Language Project
  envisioned by \newcite{abney2010universal} would be a powerful
  resource for the study of endangered languages.
%, \bluenote{keep? from
 %   the perspective of both the documentary and the computational
  %  linguist}. 
  Existing corpora are limited in the range of languages covered, in
  standardisation, or in machine-readability. In this paper we present
  SeedLing, a seed corpus for the Human Language Project. We first
  survey existing efforts to compile cross-linguistic resources, then
  describe our own approach. To build the foundation text for a
  universal corpus, we crawl and clean texts from several web sources
  that contain data from a large number of languages, and convert them
  into a standardised form consistent with the guidelines of
  \newcite{abney2011data}. The resulting corpus is more
  easily-accessible and machine-readable than any of the underlying
  data sources, and, with data from 1347 languages covering 106
  language families, represents a significant base corpus for
  researchers to draw on and add to in the future. To show the utility
  of SeedLing for cross-lingual computational research, we use our
  data in the test application of detecting similar languages.
\end{abstract}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  A broad-coverage corpus such as the Human Language Project
  envisioned by \newcite{abney2010universal} would be a powerful
  resource for the study of endangered languages.
%, \bluenote{keep? from
 %   the perspective of both the documentary and the computational
  %  linguist}. 
  Existing corpora are limited in the range of languages covered, in
  standardisation, or in machine-readability. In this paper we present
  SeedLing, a seed corpus for the Human Language Project. We first
  survey existing efforts to compile cross-linguistic resources, then
  describe our own approach. To build the foundation text for a
  universal corpus, we crawl and clean texts from several web sources
  that contain data from a large number of languages, and convert them
  into a standardised form consistent with the guidelines of
  \newcite{abney2011data}. The resulting corpus is more
  easily-accessible and machine-readable than any of the underlying
  data sources, and, with data from 1451 languages covering 105
  language families, represents a significant base corpus for
  researchers to draw on and add to in the future. To show the utility
  of SeedLing for cross-lingual computational research, we use our
  data in the test application of detecting similar languages.
\end{abstract}
